---
title: "krsken2"
author: "Vanessa Dumeaux"
date: "5/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
kdb <- "/home/shared/data/db/kraken2_microbiome"

system(capture.output(
    cat("/home/data/analysis-tools/kraken/kraken2-build --download-taxonomy -use-ftp --db", kdb)))

```

```{r}
system(capture.output(
    cat("/home/data/analysis-tools/kraken/kraken2-build --download-library plasmid -use-ftp --db", kdb)))


system(capture.output(
    cat("/home/data/analysis-tools/kraken/kraken2-build --download-library viral -use-ftp --db", kdb)))


system(capture.output(
    cat("/home/data/analysis-tools/kraken/kraken2-build --download-library protozoa -use-ftp --db", kdb)))

system(capture.output(
    cat("/home/data/analysis-tools/kraken/kraken2-build --download-library UniVec -use-ftp --db", kdb)))

system(capture.output(
    cat("/home/data/analysis-tools/kraken/kraken2-build --download-library bacteria -use-ftp --db", kdb)))

system(capture.output(
    cat("/home/data/analysis-tools/kraken/kraken2-build --download-library archaea -use-ftp --db", kdb)))

system(capture.output(
    cat("/home/data/analysis-tools/kraken/kraken2-build --download-library fungi -use-ftp --db", kdb)))

system(capture.output(
    cat("/home/data/analysis-tools/kraken/kraken2-build --download-library human -use-ftp --db", kdb)))

system(capture.output(
    cat("/home/data/analysis-tools/kraken/kraken2-build --download-library nt -use-ftp --db", kdb)))

```

### add microbiome databases

```{r}
system(capture.output(
    cat("find /home/shared/data/db/gutMicrobiome/sgbMicrobiome/representatives -name '*.fa' -print0 | \
        xargs -0 -I{} -n1 /home/data/analysis-tools/kraken/kraken2-build --add-to-library {} --db", kdb)
))
system(capture.output(
    cat("find /home/shared/data/db/gutMicrobiome -name '*.fna.gz' -print0 | \
        xargs -0 -I{} -n1 /home/data/analysis-tools/kraken/kraken2-build --add-to-library {} --db", kdb)
))

```

```{r}
system(capture.output(cat("/home/data/analysis-tools/kraken/kraken2-build --build --threads=48 --db", kdb)))

system(capture.output(cat("/home/data/analysis-tools/kraken/kraken2-inspect --db",kdb, "| head -20")))
```

```{bash}
bash ~/repositories/microbiome2brain/exp/exp01-preprocessing/kraken2-reads.sh
```

```{r}
system(capture.output(
                      cat("/home/data/analysis-tools/Bracken/bracken-build -t 48 -k 35 -l 150 -d", kdb, "-x /home/data/analysis-tools/kraken/" )
                      ))

```


# Running Kraken Reports in Bracken 

```{r}

library(R.utils)
library(BiocParallel)

source("../../src/preprocessing.R")

ncores <- 24
register(MulticoreParam(workers=ncores))

sn <- list.dirs("/home/data/refined/microbiome2brain/microbiome/01-trimmed", full.names = FALSE, recursive = FALSE)

bplapply(sn, bracken)
```


PROGRAM START TIME: 06-15-2020 01:47:13
BRACKEN SUMMARY (Kraken report: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E1A_B1/reads.kreport)
    >>> Threshold: 10 
    >>> Number of species in sample: 15987 
	  >> Number of species with reads > threshold: 7504 
	  >> Number of species with reads < threshold: 8483 
    >>> Total reads in sample: 29316385
	  >> Total reads kept at species level (reads > threshold): 16993226
	  >> Total reads discarded (species reads < threshold): 18789
	  >> Reads distributed: 7964327
	  >> Reads not distributed (eg. no species above threshold): 4737
	  >> Unclassified reads: 4335307
BRACKEN OUTPUT PRODUCED: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E1A_B1/reads.bracken
PROGRAM END TIME: 06-15-2020 01:47:21
  Bracken complete.
PROGRAM START TIME: 06-15-2020 01:47:14
BRACKEN SUMMARY (Kraken report: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E1B_B2/reads.kreport)
    >>> Threshold: 10 
    >>> Number of species in sample: 15528 
	  >> Number of species with reads > threshold: 7303 
	  >> Number of species with reads < threshold: 8225 
    >>> Total reads in sample: 25145154
	  >> Total reads kept at species level (reads > threshold): 16005416
	  >> Total reads discarded (species reads < threshold): 18478
	  >> Reads distributed: 5223673
	  >> Reads not distributed (eg. no species above threshold): 4352
	  >> Unclassified reads: 3893235
BRACKEN OUTPUT PRODUCED: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E1B_B2/reads.bracken
PROGRAM END TIME: 06-15-2020 01:47:22
  Bracken complete.
PROGRAM START TIME: 06-15-2020 01:47:14
BRACKEN SUMMARY (Kraken report: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E2A_A3/reads.kreport)
    >>> Threshold: 10 
    >>> Number of species in sample: 17473 
	  >> Number of species with reads > threshold: 7890 
	  >> Number of species with reads < threshold: 9583 
    >>> Total reads in sample: 30821677
	  >> Total reads kept at species level (reads > threshold): 17651055
	  >> Total reads discarded (species reads < threshold): 19784
	  >> Reads distributed: 8631100
	  >> Reads not distributed (eg. no species above threshold): 3672
	  >> Unclassified reads: 4516066
BRACKEN OUTPUT PRODUCED: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E2A_A3/reads.bracken
PROGRAM END TIME: 06-15-2020 01:47:22
  Bracken complete.
PROGRAM START TIME: 06-15-2020 01:47:15
BRACKEN SUMMARY (Kraken report: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E2B_A4/reads.kreport)
    >>> Threshold: 10 
    >>> Number of species in sample: 13676 
	  >> Number of species with reads > threshold: 6531 
	  >> Number of species with reads < threshold: 7145 
    >>> Total reads in sample: 12512281
	  >> Total reads kept at species level (reads > threshold): 7278937
	  >> Total reads discarded (species reads < threshold): 17100
	  >> Reads distributed: 3545067
	  >> Reads not distributed (eg. no species above threshold): 2532
	  >> Unclassified reads: 1668646
BRACKEN OUTPUT PRODUCED: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E2B_A4/reads.bracken
PROGRAM END TIME: 06-15-2020 01:47:22
  Bracken complete.
PROGRAM START TIME: 06-15-2020 01:47:15
BRACKEN SUMMARY (Kraken report: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E3A_A5/reads.kreport)
    >>> Threshold: 10 
    >>> Number of species in sample: 17167 
	  >> Number of species with reads > threshold: 7453 
	  >> Number of species with reads < threshold: 9714 
    >>> Total reads in sample: 39605876
	  >> Total reads kept at species level (reads > threshold): 23946548
	  >> Total reads discarded (species reads < threshold): 20069
	  >> Reads distributed: 10154433
	  >> Reads not distributed (eg. no species above threshold): 9090
	  >> Unclassified reads: 5475736
BRACKEN OUTPUT PRODUCED: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E3A_A5/reads.bracken
PROGRAM END TIME: 06-15-2020 01:47:23
  Bracken complete.
PROGRAM START TIME: 06-15-2020 01:47:15
BRACKEN SUMMARY (Kraken report: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E3B_A6/reads.kreport)
    >>> Threshold: 10 
    >>> Number of species in sample: 15956 
	  >> Number of species with reads > threshold: 7169 
	  >> Number of species with reads < threshold: 8787 
    >>> Total reads in sample: 35345169
	  >> Total reads kept at species level (reads > threshold): 21551109
	  >> Total reads discarded (species reads < threshold): 18327
	  >> Reads distributed: 10207733
	  >> Reads not distributed (eg. no species above threshold): 4866
	  >> Unclassified reads: 3563135
BRACKEN OUTPUT PRODUCED: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E3B_A6/reads.bracken
PROGRAM END TIME: 06-15-2020 01:47:24
  Bracken complete.
PROGRAM START TIME: 06-15-2020 01:47:16
BRACKEN SUMMARY (Kraken report: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E4A_A7/reads.kreport)
    >>> Threshold: 10 
    >>> Number of species in sample: 13745 
	  >> Number of species with reads > threshold: 5892 
	  >> Number of species with reads < threshold: 7853 
    >>> Total reads in sample: 32508399
	  >> Total reads kept at species level (reads > threshold): 20518894
	  >> Total reads discarded (species reads < threshold): 19716
	  >> Reads distributed: 10237296
	  >> Reads not distributed (eg. no species above threshold): 3288
	  >> Unclassified reads: 1729205
BRACKEN OUTPUT PRODUCED: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E4A_A7/reads.bracken
PROGRAM END TIME: 06-15-2020 01:47:24
  Bracken complete.
PROGRAM START TIME: 06-15-2020 01:47:17
BRACKEN SUMMARY (Kraken report: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E4B_A8/reads.kreport)
    >>> Threshold: 10 
    >>> Number of species in sample: 15365 
	  >> Number of species with reads > threshold: 6821 
	  >> Number of species with reads < threshold: 8544 
    >>> Total reads in sample: 44316027
	  >> Total reads kept at species level (reads > threshold): 28494524
	  >> Total reads discarded (species reads < threshold): 18783
	  >> Reads distributed: 12999308
	  >> Reads not distributed (eg. no species above threshold): 4172
	  >> Unclassified reads: 2799240
BRACKEN OUTPUT PRODUCED: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E4B_A8/reads.bracken
PROGRAM END TIME: 06-15-2020 01:47:25
  Bracken complete.
PROGRAM START TIME: 06-15-2020 01:47:17
BRACKEN SUMMARY (Kraken report: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E5A_B3/reads.kreport)
    >>> Threshold: 10 
    >>> Number of species in sample: 14129 
	  >> Number of species with reads > threshold: 6627 
	  >> Number of species with reads < threshold: 7502 
    >>> Total reads in sample: 27129895
	  >> Total reads kept at species level (reads > threshold): 16882975
	  >> Total reads discarded (species reads < threshold): 18165
	  >> Reads distributed: 6304020
	  >> Reads not distributed (eg. no species above threshold): 3495
	  >> Unclassified reads: 3921240
BRACKEN OUTPUT PRODUCED: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E5A_B3/reads.bracken
PROGRAM END TIME: 06-15-2020 01:47:25
  Bracken complete.
PROGRAM START TIME: 06-15-2020 01:47:18
BRACKEN SUMMARY (Kraken report: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E5B_B4/reads.kreport)
    >>> Threshold: 10 
    >>> Number of species in sample: 13644 
	  >> Number of species with reads > threshold: 6448 
	  >> Number of species with reads < threshold: 7196 
    >>> Total reads in sample: 29439361
	  >> Total reads kept at species level (reads > threshold): 18874712
	  >> Total reads discarded (species reads < threshold): 17816
	  >> Reads distributed: 7369578
	  >> Reads not distributed (eg. no species above threshold): 3649
	  >> Unclassified reads: 3173606
BRACKEN OUTPUT PRODUCED: /home/data/refined/microbiome2brain/microbiome/01-trimmed/E5B_B4/reads.bracken
PROGRAM END TIME: 06-15-2020 01:47:26
  Bracken complete.
PROGRAM START TIME: 06-15-2020 01:47:18
BRACKEN SUMMARY (Kraken report: /home/data/refined/microbiome2brain/microbiome/01-trimmed/F1A_A1/reads.kreport)
    >>> Threshold: 10 
    >>> Number of species in sample: 18266 
	  >> Number of species with reads > threshold: 7972 
	  >> Number of species with reads < threshold: 10294 
    >>> Total reads in sample: 30081018
	  >> Total reads kept at species level (reads > threshold): 20730642
	  >> Total reads discarded (species reads < threshold): 20935
	  >> Reads distributed: 4350114
	  >> Reads not distributed (eg. no species above threshold): 4713
	  >> Unclassified reads: 4974614
BRACKEN OUTPUT PRODUCED: /home/data/refined/microbiome2brain/microbiome/01-trimmed/F1A_A1/reads.bracken
PROGRAM END TIME: 06-15-2020 01:47:26
  Bracken complete.
PROGRAM START TIME: 06-15-2020 01:47:18
BRACKEN SUMMARY (Kraken report: /home/data/refined/microbiome2brain/microbiome/01-trimmed/F1B_A2/reads.kreport)
    >>> Threshold: 10 
    >>> Number of species in sample: 17525 
	  >> Number of species with reads > threshold: 7732 
	  >> Number of species with reads < threshold: 9793 
    >>> Total reads in sample: 24880463
	  >> Total reads kept at species level (reads > threshold): 16994177
	  >> Total reads discarded (species reads < threshold): 20251
	  >> Reads distributed: 3748913
	  >> Reads not distributed (eg. no species above threshold): 4803
	  >> Unclassified reads: 4112319
BRACKEN OUTPUT PRODUCED: /home/data/refined/microbiome2brain/microbiome/01-trimmed/F1B_A2/reads.bracken
PROGRAM END TIME: 06-15-2020 01:47:26
  Bracken complete.
PROGRAM START TIME: 06-15-2020 01:47:19
BRACKEN SUMMARY (Kraken report: /home/data/refined/microbiome2brain/microbiome/01-trimmed/F2A_B5/reads.kreport)
    >>> Threshold: 10 
    >>> Number of species in sample: 14654 
	  >> Number of species with reads > threshold: 6903 
	  >> Number of species with reads < threshold: 7751 
    >>> Total reads in sample: 30237955
	  >> Total reads kept at species level (reads > threshold): 15244345
	  >> Total reads discarded (species reads < threshold): 17600
	  >> Reads distributed: 11374175
	  >> Reads not distributed (eg. no species above threshold): 3470
	  >> Unclassified reads: 3598365
BRACKEN OUTPUT PRODUCED: /home/data/refined/microbiome2brain/microbiome/01-trimmed/F2A_B5/reads.bracken
PROGRAM END TIME: 06-15-2020 01:47:27
  Bracken complete.
PROGRAM START TIME: 06-15-2020 01:47:19
BRACKEN SUMMARY (Kraken report: /home/data/refined/microbiome2brain/microbiome/01-trimmed/F2B_B6/reads.kreport)
    >>> Threshold: 10 
    >>> Number of species in sample: 14157 
	  >> Number of species with reads > threshold: 6839 
	  >> Number of species with reads < threshold: 7318 
    >>> Total reads in sample: 27095464
	  >> Total reads kept at species level (reads > threshold): 15454471
	  >> Total reads discarded (species reads < threshold): 17465
	  >> Reads distributed: 8314827
	  >> Reads not distributed (eg. no species above threshold): 3094
	  >> Unclassified reads: 3305607
BRACKEN OUTPUT PRODUCED: /home/data/refined/microbiome2brain/microbiome/01-trimmed/F2B_B6/reads.bracken
PROGRAM END TIME: 06-15-2020 01:47:27
  Bracken complete.



### Run kraken2 on contigs
```{bash}
bash ~/repositories/microbiome2brain/exp/exp01-preprocessing/kraken2-contigs.sh
```

```{bash}
bash ~/repositories/microbiome2brain/exp01-preprocessing/krona.sh
```

